\newpage


{\vspace{-1.2cm}\normalfont\fontsize{35}{25}\bfseries
Preface}

\vspace{2cm}

Linear algebra is the fundamental introduction of algebra. However, when I first learned this course, I usually found myself confused. In this course, there are quite a number of concepts that do not seem to be useful when you first learn them, but later when I was going to use them, I've already forgotten the specific content. \\

Later I found out, if we understand what's the propose of the content we are studying, it itself will be very helpful! That's why I've decided to write this ``lecture notes''. In fact, we don't have lectures. (Maybe in the future?) But I hope this material will be helpful for the one's who are going to learn this subject, or someone who have already learnt this but usually find it confusing about ``What is linear algebra'', the latter usually conclude that this course is used for solving linear equations.(laugh)\\

I gave the small book a subtitle called ``How does it develop and why do we study this?''. For the first part, I will try my best to show you the idea of developing the concepts. And what I want to mention about the second part is that, the ``usefulness'' of mathematics usually appears in both ``mathematics'' propose and ``application'' propose. They are, in fact, somehow connected. For example, the invention of imaginary number $i$, was invented for the ``existence'' of root of a quadratic equation. And this invention is also very useful for the later application, especially in EE or some other industries. This is also true for linear algebra. I will include adequate examples to elaborate this point.\\

The content of the book mainly comes from lecture notes of two of my professors, Professor Ivan Ip and Professor YAN Min in HKUST. Thanks for their courses for me to learn this course.\\

Now let's begin!\\

%
\pagestyle{fancy}
\fancyhf{}             % 清空默认的页眉和页脚
\fancyfoot[C]{\thepage} % 设置页脚中间显示页面编号
